%{
\documentclass[11pt, twoside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex	
\usepackage{amssymb}
\usepackage{color}
\usepackage{matlab-prettifier}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{bm}
\usepackage{hyperref} 
\usepackage[section]{placeins}
\lstset{style=Matlab-editor,basicstyle=\ttfamily}

\sloppy
\definecolor{lightgray}{gray}{0.5}
\newenvironment{matlab}{\comment}{\endcomment}
\newenvironment{matlabv}{\lstlisting}{\endlstlisting}	


\title{Homework \# 3 \\ Gaussian Processes}
\author{Cody W. Eilar}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle


\begin{matlab}
%}
close all; 
clear all;

fid = fopen('output_dir/computer.tex', 'w'); 
fprintf(fid, computer); 
fclose(fid); 

fid = fopen('output_dir/matlabver.tex', 'w'); 
a = ver('matlab'); 
fprintf(fid, [a.Name ' version ' a.Version]); 
fclose(fid); 
%{
\end{matlab}

\section{Introduction} 
In the previous homework assignments we focused on processes with support vector machines
and kernels. Support vector machines are great because no \textit{a prior} assumptions
are made about the data. This then allows them to work in a myriad of cases when a
distribution about the data is not known. However, they can become computationally 
infeasible when cross validation is introduced. When dealing with large datasets, it can
take hours or days to properly compute the accuracy of a machine using cross validation 
techniques. This is where Gaussian processes are truly a useful tool. With Gaussian processes
there is no need to cross validate because the uncertainty is built into the model and it 
is therefore safe to assume with some certainty that the data in which you are trying to 
predict or fit will fall into that assumption.

In this paper we will explore how Linear Gaussian Processes work and what their advantages, 
as well as their disadvantages are for regressing and classifying stochastic processes. 
We will also look into extending Gaussian Processes to using \textit{Recursive Kernel Hilbert
Spaces}. We then look at how to optimize the hyper-parameters using exact inference. 
This then will set us up to demonstrate this theory in a few experiments that use
\textbf{MATLAB} to demonstrate a few key principles of Gaussian Processes. 
\section{Theory}
\subsection{Linear Gaussian Process}
We begin the theory section by looking into the signal model and the optimization criterion
for linear processes. As with \textit{SVM}s we are trying to solve Equation \ref{eq:linear_eq}

\begin{align}
f(x) &= \mathbf{x}^T\mathbf{w}\label{eq:linear_eq}
\end{align}

However, in the case of Gaussian Processes, we now assume that there is additive noise 
$\epsilon$ as shown in Equation \ref{eq:add_noise}
\begin{align}
y &= f(x) + \epsilon \label{eq:add_noise}
\end{align}

We take that assumption further and say that the additive noise is independent, 
identically distributed Gaussian noise with zero mean and variance $\sigma^2_n$ \cite{gauss_proc}.
\begin{align}
\epsilon \sim  \mathcal{N}(0, \sigma^2_n) \label{eq:guass_zero}
\end{align}
This now gives rise to what is known as the \textit{likelihood}. That is to say 
given the samples, $X$ and the parameters $\bf{w}$ what is the likelihood
of $\bf{y}$. Mathematically speaking, this is the same as Equation \ref{eq:likelihood}. 
\begin{align}
p(\mathbf{y}|X,\mathbf{w}) = \prod^n_{i=1}p(y_i|\mathbf{x}_i, \mathbf{w})\label{eq:likelihood}
\end{align}
The next step is to find the posterior and the marginal likelihood
\begin{align}
\text{posterior} &= \frac{\text{likelihood} \times \text{prior}}{\text{marginal likelihood}}
\end{align}

And then finally we can write the optimized model as predictive distribution:
\begin{align}
p(f_*|\mathbf{x}_*, X, \mathbf{y}) &= \int p(f_* |\mathbf{x}_*, w)p(\mathbf{w}|X, \mathbf{y})d\mathbf{w} \\
&= \mathcal{N}(\frac{1}{\sigma^2_n} \mathbf{x}_*^TA^{-1} X \mathbf{y}, \mathbf{x}_*^T A^{-1} \mathbf{x}_*)
\label{eq:predictive_dist}
\end{align}


It as at this point where we see that there is no criterion to choose for selecting an optimal model. 
In SVMs for example, we have parameters like cost that we can tweak in order to get a more accurate model. 
Here, however, there is no need for that process because we now have a distribution of where
there data could potentially lie. In Equation \ref{eq:predictive_dist} the $*$ symbol denotes unseen
 or test data and $A = \sigma^{-2}XX^T + \Sigma_p^{-1}$.

\subsection{Gaussian process in Recursive Kernel Hilbert Spaces}
The same process that is used to create the Gaussian predictive distributions can also be used
in kernel spaces. In SVMs, the \textit{kernel trick} allowed us to pass to a higher dimensional 
space easily and still have a tractable the problem. The same is also true for Gaussian processes
\cite{gauss_proc}.In order to achieve this, we create a function $\phi(x)$ which maps a D-dimensional
input vector $x$ into an N dimensional feature space. Thus our new formulation is $f(\mathbf{x}) = 
\phi(\mathbf{x})^T\mathbf{w}$. We can analyze this model the same way that was done in the linear model. 
We start again by showing the predictive distribution as show in \ref{eq:predictive_dist_hilbert}.

\begin{align}
p(\bar{f}_*|\mathbf{x}^*, \mathbf{y}, X) & \sim 
  \mathcal{N}(\mathbf{\bar{w}}^T \mathbf{x}^*, \mathbf{x}^{*T} \mathbf{A}^{-1} \mathbf{x}^*)
\label{eq:predictive_dist_hilbert}
\end{align}

Where $A = \sigma^{-2}XX^T + \Sigma_p^{-1}$ and $\mathbf{\bar{w}} = (\mathbf{XX}^T +
\sigma_n^2 \Sigma_p^{-1})^{-1} \mathbf{Xy}$. Now we can take this formulation and apply
the kernel trick to get the following equations. 

\begin{align}
\mathbf{A} &= \sigma_n^{-2} \mathbf{\Phi \Phi}^T + \Sigma_p^{-1} \\
\mathbf{\bar{w}} &= (\mathbf{\Phi \Phi}^T + \sigma^2_n \Sigma_p^{-1})^{-1} \mathbf{\Phi y}
\end{align}

Where $\Phi$ is the aggregation of columns in $\phi(\mathbf{x})$. Now we can apply the Representer 
Theorem to $\phi(x)$ as $\phi(x) \rightarrow \Sigma^{\frac{1}{2}}_p \phi(x)$ which gives a linear
combination of the transformed data. The estimators can then be rewritten as: 

\begin{align}
\Sigma_p \mathbf{\Phi \alpha} &= (\mathbf{\Phi \Phi}^T + \sigma^2_n \Sigma^{-1}_p)^{-1} \mathbf{\Phi y} \\ 
(\mathbf{\Phi \Phi}^T + \sigma^2_n \Sigma^{-1}_p) \Sigma_p \mathbf{\Phi \alpha} &= \mathbf{\Phi y}
\label{eq:representer_phi}
\end{align}

After multiplying Equation \ref{eq:representer_phi} by $\mathbf{\Phi}^T$ we can obtain the dual 
expression as show in Equation \ref{eq:dual_gauss}.

\begin{align}
\mathbf{\alpha} &= (\mathbf{\Phi}^T \Sigma_p \mathbf{\Phi} + \sigma^2_n \mathbf{I})^{-1} \mathbf{y}\\
&= (\mathbf{K} + \sigma^2_n \mathbf{I})^{-1} \mathbf{y}
\label{eq:dual_guass}
\end{align}

\subsection{Inference over the hyperparameters}

\section{Methodology and Experiments}
\subsection{Linear Gaussian Process}
\subsubsection{Experimental setup}
\subsubsection{Linear Gaussian Process for Prediction}
\subsubsection{Linear ARMA and AR(1) process noise}
\subsection{Nonlinear Gaussian Process}

\section{Discussion and Conclusion}

\bibliography{hw3.bib}
\bibliographystyle{ieeetr}

\end{document}  
%}

